Which ITG Stepcharts are Turniest?

### 2016-03-13 itg/paper.tex
fix citations
some more fixes
more fixes

{\bf Turning is exhausting.} Consider the three stream patterns shown in Figure~\ref{fig:stream}.
In the leftmost stream, the player must barely move her feet at all, never changing facing from $U$.
In the middle stream, the player faces largely $UR$, with brief $U$ facings ($\mathcal{T}=1$) as she moves between the $L$/$D$ and $U$/$R$ arrows.
Both of these streams are very easy to step without becoming fatigued.
However, the rightmost stream features many candle steps (or beer steps), changing facing from $UR$ to $UL$ and back in the span of a single measure.
ITG players hate this \cite{weirdtrick}!
However, note that even in this stream, not {\em all} of the steps have $\mathcal{T}=2$: the {\em average} turniness of this stream is only 4/3.

@@

{\bf The Least Turny Charts.}
16 charts in our collection have $\mathcal{T}=0$. Most of these are beginner charts, rated 2 or 3 feet and with no more than 100 steps. \cite{deltamax} stands out, as a 15 with 945 steps. God is that chart ever annoying.

Ignoring all beginner-difficulty and turn-free charts, the least turny stepchart is {\em Sick Dubstep Track} \cite{sickdubstep}, a 4-measure joke rated 6,482 feet, the majority of which is shown in Figure~\ref{fig:example}(a).

@@

\begin{figure*}[t]
	\begin{center}
	\begin{tabular}{cccc}
		\hspace{-2em}
		\begin{tabular}{c}
		\includegraphics[width=0.18\textwidth]{sick-dubstep-track.png} \\
		\small
		\begin{tabular}{l}
			(a) {\em Sick Dubstep Track}, $\mathcal{T}=0.26$. \\
		The least turny non-trivial \\
		chart.\\
		\\
		\end{tabular}
		\end{tabular}
		&
		\hspace{-1em}
		\begin{tabular}{c}
		\includegraphics[width=0.18\textwidth]{mr-saxobeat.png} \\
		\small
		\begin{tabular}{l}
			(b) {\em Mr. Saxobeat}, $\mathcal{T}=1.04$. \\
		Look at all those candles! \\
		No wonder it was so hard. \\
		\\
		\end{tabular}
		\end{tabular}
		&
		\hspace{-1em}
		\begin{tabular}{c}
		\includegraphics[width=0.185\textwidth]{flames.png} \\
		\small
		\begin{tabular}{l}
			(c) {\em Flames of the Sky}, $\mathcal{T}=1.04$. \\
		False-positive with foot-\\
		switches. Actually faces \\
		$UR$/$U$/$UL$ the whole time.
		\end{tabular}
		\end{tabular}
		&
		\hspace{-1em}
		\begin{tabular}{c}
		\includegraphics[width=0.18\textwidth]{do-me.png} \\
		\small
		\begin{tabular}{l}
			(d) {\em DO ME}, $\mathcal{T}=1.63$. \\
		I'll pass, thanks.\\
		\\
		\\
		\end{tabular}
		\end{tabular}
	\end{tabular}
	\end{center}
	\caption{Real-world ITG charts of varying degrees of turniness.}
	\label{fig:example}
\end{figure*}

@@

\begin{table}[t]
	\begin{center}
		\small
	\begin{tabular}{l|c|c}
		Song Pack & \# of Charts & Avg. $\mathcal{T}$ \\
		\hline
		Valex's Magical 4-Arrow Adventure 5 & 150 & 0.964 \\
		Valex's Magical 4-Arrow Adventure 4 & 170 & 0.956 \\
		Valex's Magical 4-Arrow Adventure 3 & 155 & 0.950 \\
		Valex's Magical 4-Arrow Adventure 7 & 150 & 0.937 \\
		Valex's Magical 4-Arrow Adventure 2 & 130 & 0.929 \\
		Valex's Magical 4-Arrow Adventure   & 95 & 0.919 \\
		Valex's Magical 4-Arrow Adventure 6 & 145 & 0.917 \\
		Subluminal & 16 & 0.906 \\
		The Legend of Zim 1 & 26 & 0.906 \\
		FoxyMix 5 & 30 & 0.901 \\
		FoxyMix 4 - Nuclear Overdrive & 68 & 0.899 \\ 
		Mute Sims 9 & 48 & 0.886 \\
		In The Groove Rebirth + & 147 & 0.881 \\
		StepMania 5 & 5 & 0.869 \\
		\multicolumn{3}{c}{\normalsize\dots} \\
		Stephcharts and Richarts & 81 & 0.770 \\
		Tachyon Delta & 32 & 0.769 \\
		Tachyon Alpha & 151 & 0.769 \\
		Sexuality Violation 2 & 219 & 0.766 \\
		TLOES Chapter 1 & 120 & 0.765 \\
		Fort Rapids V (12s-16s) r3 Final & 98 & 0.759 \\
		Noisiastreamz & 20 & 0.758 \\
		In The Groove & 331 & 0.754 \\
		Pendulum & 150 & 0.747 \\
		SPEEDCOOOORE 2 & 70 & 0.731 \\
		SPEEDCOOOORE & 47 & 0.709 \\
		The Legend of Zim 3 & 40 & 0.702 \\
		The Legend of Zim 4 & 44 & 0.701 \\
		The Legend of Zim 5 & 19 & 0.667 \\
	\end{tabular}
	\end{center}
	\caption{The turniest and least turny stepfile packs. Wow, great job Valex! Zim had a good thing going but has been slacking off recently.}
	\label{tab:pack}
\end{table}
\begin{table}[t]
	\begin{center}
		\small
	\begin{tabular}{l|c|c}
		Step Artist & \# of Charts & Avg. $\mathcal{T}$ \\
		\hline
		M. Emirzian & 25 & 0.981 \\
		sssmsm & 40 & 0.980 \\
		M. Calfin & 16 & 0.947 \\
		J. Berkowitz (Valex) & 1016 & 0.940 \\
		B. Dinh & 10 & 0.919 \\
		J. DeGarmo & 24 & 0.915 \\
		X. Ythar & 17 & 0.914 \\
		R. McKanna & 59 & 0.895 \\
		R. Woods & 18 & 0.892 \\
		Fraxtil & 23 & 0.880 \\
		\multicolumn{3}{c}{\normalsize\dots} \\
		I. Pyles & 175 & 0.754 \\
		P. Shanklin & 29 & 0.752 \\
		Gazebo & 355 & 0.751 \\
		zimlord & 199 & 0.750 \\
		C. Emirzian & 17 & 0.736 \\
		Mootz & 22 & 0.734 \\
		B. Abear & 27 & 0.729 \\
		Insane Steve & 41 & 0.718 \\
		Blazing & 15 & 0.696 \\
		C. Foy (\smiley) & 138 & 0.678 \\
	\end{tabular}
	\end{center}
	\caption{The turniest and least turny step authors. Only authors with 10 or more charts are included.}
	\label{tab:artist}
\end{table}

>>>>

{\bf Turning is exhausting.} Consider the three stream patterns shown in Figure~\ref{fig:stream}.
In the leftmost stream, the player must barely move her feet at all, never changing facing from $U$.
In the middle stream, the player faces largely $UR$, with brief $U$ and $R$ facings ($\mathcal{T}=1$) as she moves between the $L$/$D$ and $U$/$R$ arrows.
Both of these streams are very easy to step without becoming fatigued.
However, the rightmost stream features many candle steps (or beer steps), changing facing from $UR$ to $UL$ and back in the span of a single measure.
ITG players hate this \cite{weirdtrick}!
However, note that even in this stream, not {\em all} of the steps have $\mathcal{T}=2$: the {\em average} turniness of this stream is only 4/3.

@@

{\bf The Least Turny Charts.}
16 charts in our collection have $\mathcal{T}=0$. Most of these are beginner charts, rated 2 or 3 feet and with no more than 100 steps. $\Delta$MAX \cite{deltamax} stands out, as a 15-footer with 945 steps. God is that chart ever annoying.

Ignoring all beginner-difficulty and turn-free charts, the least turny stepchart is {\em Sick Dubstep Track} \cite{sickdubstep}, a 4-measure joke rated 6,482 feet, the majority of which is shown in Figure~\ref{fig:example}(a). It has $\mathcal{T}=0.263$.

@@

\begin{figure*}[t]
	\begin{center}
	\begin{tabular}{cccc}
		\hspace{-2em}
		\begin{tabular}{c}
		\includegraphics[width=0.18\textwidth]{sick-dubstep-track.png} \\
		\small
		\begin{tabular}{l}
			(a) {\em Sick Dubstep Track}, $\mathcal{T}=0.26$. \\
		The least turny non-trivial \\
		chart.\\
		\\
		\end{tabular}
		\end{tabular}
		&
		\hspace{-1.5em}
		\begin{tabular}{c}
		\includegraphics[width=0.18\textwidth]{mr-saxobeat.png} \\
		\small
		\begin{tabular}{l}
			(b) {\em Mr. Saxobeat}, $\mathcal{T}=1.04$. \\
		Look at all those candles! \\
		No wonder it was so hard \\
		for the author to pass.\\
		\end{tabular}
		\end{tabular}
		&
		\hspace{-1.5em}
		\begin{tabular}{c}
		\includegraphics[width=0.185\textwidth]{flames.png} \\
		\small
		\begin{tabular}{l}
			(c) {\em Flames of the Sky}, $\mathcal{T}=1.04$. \\
			False-positive $\mathcal{T}$ from foot-\\
		switches. Actually faces $R$/$UR$/\\
		$U$/$UL$/$L$ the whole time.
		\end{tabular}
		\end{tabular}
		&
		\hspace{-1.5em}
		\begin{tabular}{c}
		\includegraphics[width=0.18\textwidth]{do-me.png} \\
		\small
		\begin{tabular}{l}
			(d) {\em DO ME}, $\mathcal{T}=1.63$. \\
		I'll pass, thanks.\\
		\\
		\\
		\end{tabular}
		\end{tabular}
	\end{tabular}
	\end{center}
	\caption{Real-world ITG charts of varying degrees of turniness.}
	\label{fig:example}
\end{figure*}

@@

\begin{table}[t]
	\begin{center}
		\small
	\begin{tabular}{l|c|c}
		Song Pack & \# of Charts & Avg. $\mathcal{T}$ \\
		\hline
		Valex's Magical 4-Arrow Adventure 5 & 150 & 0.964 \\
		Valex's Magical 4-Arrow Adventure 4 & 170 & 0.956 \\
		Valex's Magical 4-Arrow Adventure 3 & 155 & 0.950 \\
		Valex's Magical 4-Arrow Adventure 7 & 150 & 0.937 \\
		Valex's Magical 4-Arrow Adventure 2 & 130 & 0.929 \\
		Valex's Magical 4-Arrow Adventure   & 95 & 0.919 \\
		Valex's Magical 4-Arrow Adventure 6 & 145 & 0.917 \\
		Subluminal & 16 & 0.906 \\
		The Legend of Zim 1 & 26 & 0.906 \\
		FoxyMix 5 & 30 & 0.901 \\
		FoxyMix 4 - Nuclear Overdrive & 68 & 0.899 \\ 
		Mute Sims 9 & 48 & 0.886 \\
		In The Groove Rebirth + & 147 & 0.881 \\
		StepMania 5 & 5 & 0.869 \\
		\multicolumn{3}{c}{\normalsize\dots} \\
		Stephcharts and Richarts & 81 & 0.770 \\
		Tachyon Delta & 32 & 0.769 \\
		Tachyon Alpha & 151 & 0.769 \\
		Sexuality Violation 2 & 219 & 0.766 \\
		TLOES Chapter 1 & 120 & 0.765 \\
		Fort Rapids V (12s-16s) r3 Final & 98 & 0.759 \\
		Noisiastreamz & 20 & 0.758 \\
		In The Groove & 331 & 0.754 \\
		Pendulum & 150 & 0.747 \\
		SPEEDCOOOORE 2 & 70 & 0.731 \\
		SPEEDCOOOORE & 47 & 0.709 \\
		The Legend of Zim 3 & 40 & 0.702 \\
		The Legend of Zim 4 & 44 & 0.701 \\
		The Legend of Zim 5 & 19 & 0.667 \\
	\end{tabular}
	\end{center}
	\caption{The turniest and least turny stepfile packs. Wow, great job Valex! Zim had a good thing going but has been slacking off recently.}
	\label{tab:pack}
\end{table}
\begin{table}[t]
	\begin{center}
		\small
	\begin{tabular}{l|c|c}
		Step Artist & \# of Charts & Avg. $\mathcal{T}$ \\
		\hline
		M. Emirzian & 25 & 0.981 \\
		sssmsm & 40 & 0.980 \\
		M. Calfin & 16 & 0.947 \\
		J. Berkowitz (Valex) & 1016 & 0.940 \\
		B. Dinh & 10 & 0.919 \\
		J. DeGarmo & 24 & 0.915 \\
		X. Ythar & 17 & 0.914 \\
		R. McKanna & 59 & 0.895 \\
		R. Woods & 18 & 0.892 \\
		Fraxtil & 23 & 0.880 \\
		\multicolumn{3}{c}{\normalsize\dots} \\
		I. Pyles & 175 & 0.754 \\
		P. Shanklin & 29 & 0.752 \\
		Gazebo & 355 & 0.751 \\
		zimlord & 199 & 0.750 \\
		C. Emirzian & 17 & 0.736 \\
		Mootz & 22 & 0.734 \\
		B. Abear & 27 & 0.729 \\
		Insane Steve & 41 & 0.718 \\
		Blazing & 15 & 0.696 \\
		C. Foy & 138 & 0.678 \\
	\end{tabular}
	\end{center}
	\caption{The turniest and least turny step authors. Only authors with 10 or more charts are included.}
	\label{tab:artist}
\end{table}
